---
layout: post
title: Visualizing Mexican Soccer Salaries with Plotly
date: 2016-01-03
author: Arturo Cardenas
categories:
- blog
- R
- web scraping
- plotly
img: logo_liga_mx.jpg
thumb: tigres_uanl.png
---

###Intro


Back in November I saw a post using ggplot2 to visualized MLS Player Salaries. I thought I would give it a try and replicate the analysis for the Mexican Soccer League using the [newly open sourced Plotly](https://plot.ly/javascript/open-source-announcement/?utm_content=buffer31954&utm_medium=social&utm_source=twitter.com&utm_campaign=buffer) but first I had to figure out where to get the data...

<!--more-->

###The tweets

This is the tweet linking to [@tejaykodali](https://twitter.com/tejaykodali) post:
<blockquote class="twitter-tweet" lang="en"><p lang="en" dir="ltr">Visualizing MLS Player Salaries with ggplot2 <a href="https://t.co/Vh5I3wWB4M">https://t.co/Vh5I3wWB4M</a> <a href="https://twitter.com/hashtag/rtodolist?src=hash">#rtodolist</a> for LigaMX</p>&mdash; Arturo Cárdenas (@arturocm) <a href="https://twitter.com/arturocm/status/672268433904574466">December 3, 2015</a></blockquote>
<script async src="//platform.twitter.com/widgets.js" charset="utf-8"></script>

And this is the tweet liking to Plotly's announce to go Open Source!
<blockquote class="twitter-tweet" lang="en"><p lang="en" dir="ltr">Plotly&#39;s JavaScript graphing library is now Open-source and Free <a href="https://t.co/6KOBKCF5f4">https://t.co/6KOBKCF5f4</a> <a href="https://twitter.com/hashtag/todolist?src=hash">#todolist</a> <a href="https://twitter.com/hashtag/visualization?src=hash">#visualization</a> <a href="https://twitter.com/hashtag/rstats?src=hash">#rstats</a></p>&mdash; Arturo Cárdenas (@arturocm) <a href="https://twitter.com/arturocm/status/672415510794002432">December 3, 2015</a></blockquote>
<script async src="//platform.twitter.com/widgets.js" charset="utf-8"></script>

###ETL process

The initial task was to find a way to obtain the salary data. I didn't had a PDF file with all the salaries (just like the original post) so I had to become resourceful and look for ways to scrap the information from [Transfermarkt](http://www.transfermarkt.es). Using the `rvest` package I was able to loop through each team from the league and obtain compile a single table with all the required data. 

```{r library_load, warning=FALSE, tidi= TRUE, message=FALSE, echo=FALSE}
library(rvest)
library(XML)
library(data.table)
library(plotly)
library(RColorBrewer)
library(dplyr)
library(DT)
```

```{r rvest, warning=FALSE, tidi= TRUE, message=FALSE}
transfermarkt <- function (x = dara.url){
  # Create data.table for team loops
  team.value <- data.table()
  for (j in 1:length(x)){
    # Initial pull with all the liga MX names-url information
    all.ligamx <- read_html(paste0(url.t,x[j])) %>%
      html_nodes("table.items") %>%
      .[1]
    # Get teams url
    team.url <- html_nodes(all.ligamx, "a") %>%
      html_attr("href")
    team.url <- unique(team.url[team.url %like% "startseite"])
    # Get teams names
    names <- html_table(all.ligamx, fill = TRUE) %>%
      as.data.frame() %>%
      .[-1,"Club"]
    # Create data.frame with url-names
    db <- cbind.data.frame(team.url, names)

    # Get each player´s cost by team
    for (i in 1:nrow(db)){
      temp <- read_html(paste0(url.t,db[i,1])) %>%
        html_nodes("table.items") %>%
        .[1] %>%
        html_table(fill = TRUE) %>%
        as.data.frame()
      temp$team <- db[i,2]
      team.value <- rbind.data.frame(team.value, temp)
    }
  }
  # Update values
  headers <- c("number","x1","x2","x3","position","name","age","x7","value","team")
  names(team.value) <- headers
  team.value <- subset(team.value,
                       select = c("number","position","name","age","value","team"))
  return(team.value)
}
```


The salaries raw data has extra text and its formated as string. Using the self defined `cleansing` formula the values are 1) transformed to number values, and 2) converted from euros to USD at a 1.09 rate. 


```{r data_cleaning, warning=FALSE, tidi= TRUE, message=FALSE}
cleansing <- function(data = team.value){
  # Filter out all NA and edit strings to numeric values
  draft <- data[!(is.na(data$value) | is.na(data$team)),]
  draft$value.adj <- draft$value
  draft$value.adj <- gsub("miles €","000", draft$value.adj)
  draft$value.adj <- gsub("mill. €","0 000", draft$value.adj)
  draft$value.adj <- gsub(",","", draft$value.adj)
  draft$value.adj <- gsub(" ","", draft$value.adj)
  draft$value.adj <- as.numeric(draft$value.adj)
  draft$value.adj <- draft$value.adj * 1.09 / 1000000 #from euros to MM dlls
  draft <- draft[!(is.na(draft$value.adj)),]
  draft <- arrange(draft, desc(value.adj))
  return(draft)
}
```


In order to have more control on the colors asigned in the plot we are defining the `color.column` formula to take advante of the `RColorBrewer`package and pre-define the color scale to be used in the final plot. 


```{r colorbrew, warning=FALSE, tidi= TRUE, message=FALSE}
color.column <- function(x = data, team_rank = 0){
  draft <- arrange(x, value.adj)
  set.color <- cbind.data.frame(value.adj = unique(draft$value.adj),
                                color.set =colorRampPalette(
                                  brewer.pal(9,"Spectral"))(
                                    length(unique(draft$value.adj))))
  draft <- full_join(draft, as.data.table(set.color), by = "value.adj")
  b <- aggregate(value.adj ~ team, data=draft, FUN = sum)
  names(b) <- c("team","total")
  b <- b[order(-b$total),]
  if (team_rank == 0){
    b <- b}
  else{
    b <- b[1:team_rank,]
  }
  draft <- inner_join(draft, b, by = "team")
  draft <- draft[order(draft$value.adj),]
  return(draft)
}
```


**Finally** we are defining the parameters to be used in the **Plotly** bar chart!


```{r plotly, warning=FALSE, tidi= TRUE, message=FALSE}
league.plot <- function(x = data, Plot.Title = "League"){
  p <- plot_ly(data = x, y = team, x = value.adj,
               marker = list(color = color.set,
                             line= list(color='white', width=.5 )),
               opacity = .9, group = total, type = "bar",
               showlegend = FALSE, bargap = 0.5, text = name,
               orientation = "h")

  f1 <- list(family = "Arial, sans-serif", size = 18, color = "lightgrey")
  f2 <- list(family = "Old Standard TT, serif", size = 12, color = "black")
  x <- list(title = "USD Million", titlefont = f1, showticklabels = TRUE,
    tickfont = f2, exponentformat = "e", showexponent = "All")
  y <- list(title = "TEAM", titlefont = f1, showticklabels = TRUE,
    tickfont = f2, exponentformat = "e", showexponent = "All")
  m = list(l = 150, r = 20, b = 50, t = 50, pad = 4)
  layout(p, title=Plot.Title, hovermode = "closest", barmode = "stack", 
         xaxis = x, yaxis = y, margin = m)
}
```


Now lets go down to business! Firts we need to set the url that will direct the web scraper to the Mexican Soccer League initial page:


```{r ligamx_links, warning=FALSE, tidi= TRUE, message=FALSE}
# Set the initial url to pull the data
url.t <- "http://www.transfermarkt.es"
# Mexican League link
url.mx <- "/liga-mx-clausura/marktwerteverein/wettbewerb/MEX1/" #mexico
```


And then we´ll just run all the previously defined formulas to end up with an interactive chart:


```{r liga_mx, warning=FALSE, tidi= TRUE, message=FALSE}
team.value <- transfermarkt(url.mx)
draft <- cleansing(team.value)
final.draft <- color.column(draft, team_rank = 0)
final.draft <- arrange(final.draft, (total), (value.adj))
league.plot(final.draft, "Liga MX Teams")
```


###Going the extra mile


Now, if we want to see the same visualization taking into account all the American Continent teams, we can create a vector containing all the urls to all the leagues and pass it through the formulas.


```{r other_links, warning=FALSE, tidi= TRUE, message=FALSE}
#  Links for all the soccer leagues from america available
url.us <- "/major-league-soccer/marktwerteverein/wettbewerb/MLS1" #US
url.br <- "/campeonato-brasileiro-serie-a/marktwerteverein/wettbewerb/BRA1" #Brazil
url.co <- "/liga-postobon-i/marktwerteverein/wettbewerb/COLP" #Colombia
url.ec <- "/serie-a-primera-etapa/marktwerteverein/wettbewerb/ECPE" #Ecuador
url.ur <- "/primera-division-apertura/marktwerteverein/wettbewerb/URU1" #Uruguay
url.ch <- "/primera-division-apertura/marktwerteverein/wettbewerb/CHLA" #Chile
url.cr <- "/primera-division-invierno/marktwerteverein/wettbewerb/CRPD" #C Rica
url.pe <- "/torneo-descentralizado-clausura-entscheidungsspiel/marktwerteverein/wettbewerb/PERD" #Peru
url.ar <- "/primera-division/marktwerteverein/wettbewerb/AR1N" #Argentina
# Add all the american leagues
url.all<- c(url.us, url.br, url.ar, url.co, 
            url.ec, url.ur, url.ch, url.mx)
# Here is the link in case you want to add "La Liga" from Spain
url.es <- "/primera-division/marktwerteverein/wettbewerb/ES1"
```


For visualization porpuses we are limiting the list to the 20 most expensive soccer teams in the continent.


```{r america, warning=FALSE, tidi= TRUE, message=FALSE}
team.value <- transfermarkt(url.all)
draft <- cleansing(team.value)
final.draft <- color.column(draft, team_rank = 20)
final.draft <- arrange(final.draft, (total), (value.adj))
league.plot(final.draft, "Most Expensive Soccer teams from America")
```


And finally - since we already have the data loaded - here is the list of the 20 players with the highest salaries in the continent:


```{r players, warning=FALSE, tidi= TRUE, message=FALSE}
draft <- cleansing(team.value)
final.draft <- color.column(draft)
final.table <- select(final.draft, number, name, position, team, USD = value.adj)
final.table <- arrange(final.table, desc(USD))
final.table[1:20,]
```

